A Survey of Physical Design Techniques of Information Systems

### 2010-07-10 trunk/cs848-db/Paper.tex

In a large database system, there are usually a few complex queries that require the joining of many large tables. If these queries are frequently run, then the
database performance is likely to suffer, and users have to wait for a long time to get the results back. If we know these queries beforehand, it makes sense to
simply store the results of the queries on disk instead of recomputing them each time. This is exactly what materialized views aim to do. By precalculating the
results of a complex query, and storing them in a table on disk, the new table with the results is very likely to be much smaller than the original tables
decreasing I/O access costs and thus increasing performance~\cite{lightstone2007physical, chaudhuri1998overview}. Once a materialized view exists, the user can
either explicitly query the materialized view, or enter the original query, and have the query optimizer rewrite the query to use the materialized views instead
of the original tables. Many query optimization techniques have relied on rewriting queries using views (E.g.~\cite{levy1995answering, gupta1995aggregate,
goldstein2001optimizing,
abiteboul1998complexity}), but discussing the details of the rewriting is beyond the scope of this survey.

@@

cite~\cite{dewitt1984implementation}.

\subsubsection*{B Trees}
Although B Trees, explained in the previous section, are originally designed for disk based database systems, they were found to also be useful in
MMDB~\cite{lehman1986study}. In disk based systems, the B+ Tree is more used than the regular B Tree. However, in main memory databases, the B Tree would be
more suitable since there is no gain from keeping all the data in the leaves. In a main memory system, this would only waste space without improving
performance. In a B or B+ Tree, multiple keys are used to search within a node. This means that if the node fits in a cache line, this cache load can satisfy
more than more comparison leading to better cache utilization~\cite{rao1999cache}. Therefore, B Trees in general have good cache behavior.

@@

Binary Search Trees have poor cache behavior when the array is much bigger than the cache. Rao and Ross~\cite{rao1999cache} propose the Cache Sensitive Search
Tree (CSS-Tree) to improve this cache behavior. To do that, the search tree is stored as a directory in an array. The number of keys per node is chosen such
that the whole node fits in a cache line. This improves local searching within a node to just one cache miss. They also hard-code the traversal within a node
such that finding the next node happens in $log_{m+1}n$ times rather than $log_{2}n$ times in a binary search tree. However, CSS Trees are only suitable for
read environments since they eliminate nearly all child pointers, thus removing the support for incremental update.

\subsubsection*{J+-trees and pJ+-trees}
\cite{luan2009prefetching}

>>>>

In a large database system, there are usually a few complex queries that require the joining of many large tables. If these queries are frequently run, then the
database performance is likely to suffer, and users have to wait for a long time to get the results back. If we know these queries beforehand, it makes sense to
simply store the results of the queries on disk instead of recomputing them each time. Oracle 8i (SOURCE) introduced materialized views to fit this purpose. By
precalculating the
results of a complex query, and storing them in a table on disk, the new table with the results is very likely to be much smaller than the original tables
decreasing I/O access costs and thus increasing performance~\cite{lightstone2007physical, chaudhuri1998overview}. Once a materialized view exists, the user can
either explicitly query the materialized view, or enter the original query, and have the query optimizer rewrite the query to use the materialized views instead
of the original tables. Many query optimization techniques have relied on rewriting queries using views (E.g.~\cite{levy1995answering, gupta1995aggregate,
goldstein2001optimizing,
abiteboul1998complexity}), but discussing the details of the rewriting is beyond the scope of this survey.

@@

In this section, we talk about three families of index structures used in MMDB: B Trees, T Trees, and Binary Search Trees. There are more recent index
structures developed for MMDB such as J+ Trees~\cite{luan2009prefetching}, but which have not gained much popularity, and so we do not discuss them in our
survey. Table~\ref{tab:mmdbindexsumm} summarizes the features of the different index structures discussed in this section with respect to MMDB.

\begin{table}[t!]
\centering
\scalebox{0.7}{
\begin{tabular}{p{2cm}|p{6cm}|p{6cm}}
\hline
 & \textbf{Features suitable for MMDB} & \textbf{Cache Consciousness}\\
\hline\hline
B Trees &
\begin{itemize}\item Good storage utilization
\item Reasonably quick searching
\item Fast updating
\end{itemize} 
& \begin{itemize}\item Reasonable cache behavior if node fits in cache line\end{itemize}\\
\hline
B+ Trees & 
\begin{itemize}
\item Wastes space since all data is stored in the leaves
\item Reasonably quick searching
\item Fast updating
\end{itemize} &
\begin{itemize}
\item Reasonable cache behavior if node fits in cache line\end{itemize}\\
\hline
CSB+ Trees & \begin{itemize}
              \item Same features of B+ Trees
             \end{itemize} &
\begin{itemize}
 \item Good cache performance due to improved locality \& lower tree height
\end{itemize}\\
\hline
pB+ Trees & \begin{itemize}
              \item Same features of B+ Trees
             \end{itemize} &
\begin{itemize}
 \item Improves cache miss performance by prefetching more data in each cache miss
\item Node size is also increased leading to the same benefits of CSB+ Trees
\end{itemize}\\
\hline
T Tree &  \begin{itemize}
              \item Contain pointers to data values instead of the values themselves leading to better space utilization
\item Node size is bigger leading to previous mentioned advantages
             \end{itemize} &
\begin{itemize}
 \item Cache behavior was not considered at time of design
\end{itemize}\\
\hline
CST & \begin{itemize}
       \item Fast traversal in $log_{m+1}n$
      \end{itemize} &
\begin{itemize}
 \item Good cache behavior as number of key nodes per line is chosen according to cache line miss.
\end{itemize}\\
%\hline
%J+ Tree & \begin{itemize}
%       \item 
%      \end{itemize} &
%\begin{itemize}
% \item 
%\end{itemize}\\
%\hline
%pJ+ Tree & \begin{itemize}
%       \item Fast traversal in $log_{m+1}n$
%      \end{itemize} &
%\begin{itemize}
% \item
%\end{itemize}\\
%\hline
\end{tabular}
}
\caption{Summary of index structures for Main Memory Databases (MMDB)}
\label{tab:mmdbindexsumm}
\end{table}

%cite~\cite{dewitt1984implementation}.

\subsubsection*{B Trees}
Although B Trees, explained in the previous section, are originally designed for disk based database systems, they were found to also be useful in
MMDB~\cite{lehman1986study}. In disk based systems, the B+ Tree is more used than the regular B Tree. However, in main memory databases, the B Tree would be
more suitable since there is no gain from keeping all the datGood storage utilizationa in the leaves. In a main memory system, this would only waste space
without improving
performance. In a B or B+ Tree, multiple keys are used to search within a node. This means that if the node fits in a cache line, this cache load can satisfy
more than more comparison leading to better cache utilization~\cite{rao1999cache}. Therefore, B and B+ Trees in general have reasonable cache behavior.

@@

Binary Search Trees have poor cache behavior when the array is much bigger than the cache. Rao and Ross~\cite{rao1999cache} propose the Cache Sensitive Search
Tree (CSS-Tree) to improve this cache behavior. To do that, the search tree is stored as a directory in an array. The number of keys per node is chosen such
that the whole node fits in a cache line. This improves local searching within a node to just one cache miss. They also hard-code the traversal within a node
such that finding the next node happens in $log_{m+1}n$ times rather than $log_{2}n$ times in a binary search tree where $m$ is the number of keys per node.
However, CSS Trees are only suitable for read environments since they eliminate nearly all child pointers, thus removing the support for incremental update.

\subsubsection{Materialized Views}
From our literature survey, it seems that materialized views are not used in MMDB. This makes sense since the problems materialized views try to address are
not present in MMDB. That is, the cost of computing complicated join queries is much less in MMDB due to the low cost of accessing the data in memory.
Accordingly, the maintenance cost of materialized views in MMDB will outweigh their minimal benefit.

\subsubsection{Partitioning}
